---
title: "Penguins Plot"
author: "Sam"
date: "11/29/2021"
output:
  pdf_document: default
  html_document: default
---

## Setting up my environment
Notes: setting up my R environment by loading the 'tidyverse' and 'palmerpenguins' packages

```{r loading packages}
library(tidyverse)
library(palmerpenguins)
```

## Visualizations 
Here we will go through a series of visualizations

### Flipper and body mass in purple
Here, we plot flipper length against body mass

```{r visualize flipper and body mass in purple, message=FALSE, warning=FALSE}
ggplot(penguins, aes(flipper_length_mm, body_mass_g))+
  geom_point(color = "purple")
```


### Flipper and body mass by species
Here, we plot flipper length against body mass and look at the breakdown by species

```{r species, warning=FALSE}
ggplot(penguins, aes(flipper_length_mm, body_mass_g))+
  geom_point(aes(shape=species, color = species))
```

### Flipper and body mass by species and sex
Here, we plot flipper length agaisnt body mass and look at the breakdown by species and sex

```{r sex and species, message=FALSE, warning=FALSE}
ggplot(penguins,aes(flipper_length_mm, body_mass_g))+
  geom_point(aes(color=species,
                 shape=species))+
  facet_wrap(~sex)
```

## Penguin Plots

You can also embed plots, for example:

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r full picture, warning=FALSE}
library(palmerpenguins)
library(tidyverse)
library(lubridate)

penguins %>% 
  ggplot(aes(flipper_length_mm, body_mass_g)) + 
  geom_point(aes(shape = species, color = species))+
  geom_smooth(se=FALSE)+
  facet_wrap(~species)

```

